---
title: "Statistics_Final"
author: "Jasleen Kaur"
date: "03/04/2020"
output:
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
options(warn=-1)
library(tidyverse)
```
# What percentage of patients miss their medical appoitment and why?
# 7.1-Data Familiarity
```{r}
# read the table
med_data <-  read_csv("KaggleV2-May-2016.csv")
```

**7.1.1 Data Visualize**
```{r}
str(med_data)
```

**7.1.2 Data Clean**
```{r}
# changing columns name
names(med_data)<-  c('patient_id','appointment_id','gender','schedule_day',
                  'appointment_day','age',
                   'neighborhood','scholarship','hypertension','diabetes',
                  'alcoholism','handicap',
                   'sms_received','no_show_status')

# In data description no_show:'No' means patients showed up and 'yes' 
#means patients dint show up . Hence giving clear names
med_data$no_show_status[med_data$no_show_status == 'No'] <- 'Showed up'
med_data$no_show_status[med_data$no_show_status == 'Yes'] <- 'not showed up'
med_data$no_show_status <- as.factor(med_data$no_show_status)

# replace names for gender levels
med_data$gender[med_data$gender == 'M'] <- "Male"
med_data$gender[med_data$gender == 'F'] <- "Female"

med_data$day <- weekdays(as.Date(med_data$appointment_day))
```

```{r fig.height=7,fig.width=7}
# Check missing values
na_count <-sapply(med_data, function(y) sum(length(which(is.na(y)))))
na_count <- data.frame(na_count)
na_count$name<-rownames(na_count)

# Visualize NA values in data frame
p<- ggplot(data=na_count, aes(x=name,y=na_count,fill=na_count))+
  geom_bar(stat="identity")+
  theme_minimal()
p + coord_flip()
```

no NA values found in dataset

```{r}
# coverting some columns to factors
med_data <- mutate_at(med_data, vars('gender',
                   'neighborhood',
                   'scholarship',
                   'hypertension',
                   'diabetes',
                   'alcoholism',
                   'handicap',
                   'sms_received'), as.factor)
```
```{r}
# checking summary 
summary(med_data)
```

```{r}
# removing age = -1
med_data<-med_data[!(med_data$age<0),]
summary(med_data$age)
```

# 7.2-Exploratory Analysis

**7.2.1-Age vs no show status variable**
```{r}
# Visualize age vs no show status variable 
library(ggplot2)

ggplot(med_data, aes(x = no_show_status, y = age, fill = no_show_status))+
    geom_boxplot()+ 
    ggtitle("Age vs No Show Status Boxplot")+
    theme(plot.title = element_text(hjust = 0.5))
```

- Number of patients showing up is less for all age groups
- From age vs no show status histogram proportion of showed up patients are higher 
  in the age range 60 to 80 than patient age under 40
- Mean age of patients not showing up for treatment is less than mean age of patients who showed up for treatment
```{r}
# Exact mean value statistics of ages for show/no show
dplyr::select(med_data,age,no_show_status)%>% 
    group_by(no_show_status) %>% 
    summarise(age_mean = mean(age))
```
- This signifies there is difference in age of patients showing up and not showing up 
```{r}
# COnducting test for significance of age factor;as age is numeric and 
# no_show_status is a binary factor; 
# independent 2 group t-test
# H0: True difference in means of show age and no show age = zero; age factor 
# is not significant 
# in determining show,no show factor in medical appoitnments
t.test(med_data$age ~ med_data$no_show_status)  
```
- There is significant difference in mean age of showed and not showed up patients suggesting age is
significant factor in determining medical appointment show status.

**7.2.2-Gender vs No show status Variable**
```{r}
# total patient table
cat("PATIENT TABLE")
table(med_data$no_show_status)
cat("\n\n\n","GENDER TABLE\n")
# gender table
table(med_data$gender, med_data$no_show_status)
cat("\n\n\n","GENDER PROPORTION TABLE\n")

# proportion table for gender
prop.table(table(med_data$gender, med_data$no_show_status), margin = 1)
```

- Nearly 25% of patients dont show for scheduled medical appointment
- It seems no significant difference in gender showing or not showing for medical appointment.
```{r}
# Statistical test for significance of gender
# H0: gender factor is not significant in predicting show/no show factor
chisq.test(table(med_data$gender,med_data$no_show_status), correct = FALSE)# without continuity correction
```
- From the Pearson's chi squared test p-value is higher than alpha value(~0.05) suggesting gender 
factor is not significant in predicting show no show of patients.

**7.2.3-Handicap vs no show status variable**

```{r}
# Visulizing proportion of show/no-show for each level of handicap
ggplot(med_data)+geom_bar(aes( handicap, fill = no_show_status), position = position_fill())+
   ylab('Proportion')
```
- Proportion graph for handicap variabl esuggests show/no-show factor is not really dependent on it . 
Furthur we can verify it using chi squared test
```{r}
# Pearson Chi squared test: H0- handicap factor is not significant in determining 
# show/no-show status for medical appointment
chisq.test(table(med_data$handicap, med_data$no_show_status), correct = FALSE) #handicap
```
-pvalue for handicap factor is more than alpha(~0.05) suggesting handicap is not significant in determining target factor.

**7.2.4- EDA on other factor variables**
```{r fig.height=15,fig.width=15}
library(gridExtra)
# Visualizing significance of other factors for show/no show target factor
g1 <- ggplot(med_data)+geom_bar(aes(scholarship, fill = no_show_status))
g2 <- ggplot(med_data)+geom_bar(aes(hypertension, fill =no_show_status))
g3 <- ggplot(med_data)+geom_bar(aes(diabetes, fill = no_show_status))
g4 <- ggplot(med_data)+geom_bar(aes (alcoholism, fill =  no_show_status))
g5 <- ggplot(med_data)+geom_bar(aes(sms_received, fill = no_show_status))

grid.arrange(g1,g2,g3,g4,g5, nrow = 3)
```
- The levels for each factor variables are not clearly depicting the proportion of 
show/no-show so finding their proportions
```{r fig.height=15,fig.width=15}
gg1 <- ggplot(med_data)+geom_bar(aes(scholarship, fill = no_show_status)
                                 , position = position_fill())+ ylab('Proportion')
gg2 <- ggplot(med_data)+geom_bar(aes(hypertension, fill =no_show_status),
                                 position = position_fill())+ylab('Proportion')
gg3 <- ggplot(med_data)+geom_bar(aes(diabetes, fill = no_show_status), 
                                 position = position_fill())+ ylab('Proportion')
gg4 <- ggplot(med_data)+geom_bar(aes (alcoholism, fill =  no_show_status),
                                 position = position_fill())+ylab('Proportion')
gg5 <- ggplot(med_data)+geom_bar(aes(sms_received, fill = no_show_status), 
                                 position = position_fill())+ ylab('Proportion')

grid.arrange(gg1,gg2,gg3,gg4,gg5, nrow = 3)
```

- From these plots except alcholic factor all others seems to have significant effect on show/no show status. We can determine significance using chi squared for each one of them
```{r}
# H0: Factor is significant in determining show/no-show for patient
chisq.test(table(med_data$scholarship, med_data$no_show_status), correct = FALSE) # scholarship
chisq.test(table(med_data$hypertension, med_data$no_show_status), correct = FALSE) # hypertension
chisq.test(table(med_data$diabetes, med_data$no_show_status), correct = FALSE) # diabetes
chisq.test(table(med_data$alcoholism, med_data$no_show_status), correct = FALSE) #alcoholism
chisq.test(table(med_data$sms_received, med_data$no_show_status), correct = FALSE) # sms_received
```

- p-values from chi squared test suggest that cholarship, hypertension, diabetes, sms_received group are significant in determining show/no-show for medical appointment and
as proportion graphs suggested alcohol factor is not significant 

**7.2.5- EDA on appointment day vs No Show**
```{r}
ggplot(med_data)+geom_bar(aes(day, fill = day))+
     ggtitle("Number of Appointment")+
    ylab('Count')+
    xlab('Day')+
    theme(plot.title = element_text(hjust = 0.5))
```
-Number of appointment differ across week. Some day like Wednesday and Tuesday make more appointment than other. Statistics give exact information below.
```{r}
library(lubridate)
# make days column, with label true
med_data$date <- as.Date(med_data$appointment_day)
med_data$days <- wday(med_data$date, label=TRUE)
```


```{r}
# days column 
table(med_data$days, med_data$no_show_status)
```
- Weekends tend to have lesser number of appointments and weekdays have more . Tuesday and wednesday are busy days mostly. Day factor seem to be significant for determining show no show factor. 
- Saturday has highest number of no shows. Seems like weekend appointment tend to cancel more
```{r}
# H0: weekday is significant in determining target factor
chisq.test(table(med_data$no_show_status,med_data$day))
```
-Since p value is significantly low suggesting showing up in appointment day is dependent on which day the appointment is.

**7.2.6- EDA on Neighborhood vs No Show**
```{r}
df_neighbor <- data.frame(table(med_data$neighborhood, med_data$no_show_status))
names(df_neighbor) <- c("neighborhood", "no_show", 'Count')
```

```{r fig.height=15,fig.width=15}
# visualization on neighbourhood
gg_neigh<-ggplot(df_neighbor)+
geom_bar(aes(x = reorder(neighborhood, -Count), y = Count, fill = no_show), 
    stat = 'identity')+
    theme(axis.text.x = element_text(size= 12, angle = 90, hjust = 1))+
    ggtitle("Neighborhood vs No Show")+
    ylab('Count')+
    xlab('Neighborhood')+
    theme(plot.title = element_text(hjust = 0.5, size = 24))+
    theme(axis.title.y = element_text(size =18))+ 
    theme(axis.title.x = element_text(size =18))
gg_neigh+coord_flip()
```
There is no clear analysis on how neighbourhood is affecting appointment no show status so plotting proportions
```{r fig.height=15,fig.width=15}
# proportion 
gg_neighbour<-ggplot(med_data)+
    geom_bar(aes(x = neighborhood, fill = no_show_status), position = position_fill())+
    theme(axis.text.x = element_text(size= 12, angle = 90, hjust = 1))+
      ggtitle("Neighborhood vs No Show")+
    ylab('Proportion')+
    xlab('Neighborhood')+
    theme(plot.title = element_text(hjust = 0.5))+
    theme(plot.title = element_text(hjust = 0.5, size = 24))+
    theme(axis.title.y = element_text(size =18))+ 
    theme(axis.title.x = element_text(size =18))
gg_neighbour+coord_flip()
```
-ILHAS OCEANICAS DE TRINDADE neighbourhood has maximum no show counts of patients followed by SANTOS DUMONT area 

# 7.3- Model Selection and Predictor Significance

```{r}
# considering important factors 
med_data_2 <- dplyr::select(med_data, age, gender, scholarship, hypertension, diabetes,
                            alcoholism, handicap,sms_received, day, no_show_status)
med_data_2 <- mutate_at(med_data_2, vars(day), as.factor)
str(med_data_2)
```

```{r}
library(faraway)
log_model_1 <- glm(no_show_status ~ . ,family = binomial(link = 'logit'),  data = med_data_2 )
sumary(log_model_1)
```
- We can observe negative coefficients for scholarship true,diabetes,alcoholism,handicap,sms_received,weekend day saturday indicating number of no shows 
of patients increases if these are true for patient. While for male gender,weekdays,hypertension and age factors number of no shows are lesser .
```{r}
exp(coef(log_model_1))

```

- We can interpret coefficients for:
(a) saturday - 0.8653094 ODDS- the odds of showing up for appointment reduces by 13.4% on saturday keeping all other factors constant like age,gender,
scholarship yes or no,hypertension,diabetes,alcoholism etc
(b)hypertension-1.0695011 ODDS- the odds of showing up for appointment increases by 6.9% if patient has hypertension keeping all other factors fixed.
Seems reasonable if patient already has some medical ailment , he tends to show up for appointment.

```{r}
# we can even test significance of individual predcitor using drop1 fucntion
drop1(log_model_1,test = "Chi")
```

- Gender and handicap p-values are significantly higher indicating we can drop them from our model as they are not significant
```{r}
med_data_3 <- dplyr::select(med_data, age, scholarship, hypertension, diabetes,
                            alcoholism, sms_received, day, no_show_status)
log_model_2 <- glm(no_show_status ~ . ,family = binomial(link = 'logit'),  data = med_data_3 )
sumary(log_model_2)
```

# 7.4- Logistic Regression with training and test data set
```{r}
library(caTools)
set.seed(100)
indices = sample.split(med_data_3$no_show_status, SplitRatio = 0.7)
train = med_data_3[indices,]
test = med_data_3[!(indices),]

```

```{r}
logit_model_1 <- glm(no_show_status ~ . ,  data = train, family =binomial(link = 'logit') )
# Stepwise selection
library("MASS")
model_2<- stepAIC(logit_model_1, direction="both")
vif(logit_model_1)
```

- vif > 10 suggests inflation of factor in this case weekdays are getting inflated
```{r}
# Model from AIC obtained 
logit_model_2 <- glm(no_show_status ~ age + scholarship + hypertension + alcoholism + 
    sms_received + day ,  data = train, family =binomial(link = 'logit') )
sumary(logit_model_2)
```
- hypertension and alcoholism have significantly higher p-values suggesting they are not significant in determining show/no-show status of appointment.

```{r}
# Removing alcholism and hypertension
logit_model_3 <- glm(no_show_status ~ age + scholarship + 
    sms_received + day ,  data = train, family =binomial(link = 'logit') )
sumary(logit_model_3)
```
- Most of the p-values are significant suggesting possible significance in model building

# 7.5- Model Evaluation
```{r}
#predicted probabilities of appointment miss for test data
test_predicted = predict(logit_model_3, type = "response", newdata = test)
summary(test_predicted)
pred_test <-  ifelse(test_predicted>0.5,"No show","show")
tab <- table(predicted = pred_test, actual = test$no_show_status)
tab
```
Model is not like as we expected because there is class imbalance in between showed up and not showed up group.
There are 80 % patients who show up and 20 % those who don’t show up for appointment
```{r}
# observe residuals
qqnorm(residuals(logit_model_3))
```
